Real-Time Application of Activity Classification from Body-Worn Accelerometers

### 2016-06-05 notes/paper/paper.tex
Added ROC curve image to the paper for walking vs. jogging
Added biking ROC curve to the paper.

\documentclass[journal]{IEEEtran}
\usepackage{blindtext}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{pinyin}
\usepackage{natbib}
\usepackage{notoccite}
\usepackage{amssymb}
\usepackage{textcomp}
%
\begin{document}
%
% paper title
\title{Real-Time Application of Activity Classification from Body-Worn Accelerometers}
%
\author{Craig~Euler, C.T. Lin, Melissa Flores, et al? (Ruben, Brian, Sydney, Steve?)}
%
% make the title area
\maketitle
%
\begin{abstract}
Body worn sensors have been used in the market for many years now.
Many of which are used for activity monitoring or fitness.
In this paper, we make use of a matched-filtering algorithm for real-time activity classification from data obtained with three-axis accelerometers worn by the user.
We achieve invariance to sensor orientation with dimensionality reduction.
We make use of an instance-based learning algorithm to train the device to learn the individual's motion patterns and store that information as activity filters for our matched filter.
We improve computational efficiency with dimensionality reduction, temporal decomposition, and pre-processing of our filters for later use.
An assessment of our computational costs are provided as processing flop counts for the real-time specific parts of the algorithms.
\end{abstract}
%
\section{Introduction}
Various methods for activity classification exist.
Custom dicision tree, automatically generated decision tree, and artificial neural networks \cite{parkka_ermes_korpipaa_mantyjarvi_peltola_korhonen_2006} have been explored as well as other frequency based classification methods \cite{sharma_purwar_lee_lee_chung_2008}.
In this paper, we employ the use of a matched-filter method to identify an individual’s activity in real-time.
The matched-filtering method with use of thresholding and comparing the correlations of various filters is a simple method to use and have been studied in the past \cite{giannakis_tsatsanis_1990}.
Computational resources are limited on micro-processors so we've designed our algorithms to be as efficient as possible.
Having an a priori knowledge of the activity the individual is performing is useful for many current fitness algorithms.
These devices (Fit-Bit, Apple watch, Moov, Jawbone, etc.) usually have the individual input the activity they are performing.
%
\section{Activity Classification via Matched Filtering}
The matched-filter method is commonly used for extracting a signal out of noisy data.
We employ this method for identifying a user’s activity from data generated by body-worn accelerometers.
Because the matched-filter method requires use of an existing reference template to identify the desired signal, we must develop our template from a training set.
We will discuss how we develop our set of templates for use in our matched-filter.
%
\section{Dimensionality Reduction}
We manage to reduce our computational resources and remove our dependence on sensor orientation at the same time through a process of dimensionality reduction.
We assume that over a short period of time a user’s limb movement is constrained mostly to a two-dimensional plane.
By identifying the plane of motion, we then project the acceleration data onto this plane and re-define our $x$ and $y$ axis accordingly.
By this assumption, activities that are mostly confined to two dimensions will improve our signal identification.

@@

The training module, like the pre-processing module, does not need to be in real time.
This algorithm is the most demanding on computational resources and is rarely ran so this can be performed on a smart phone or tablet.
The purpose of the training module is to derive the reference signal to be used on the micro-controller for real-time processing.
The training process uses the same algorithm kernels as the activity classification.
%
\section{Results}
%
ROC curves, timing results?, flop counts.
%
\section{Conclusion}
conclusion text.
%
\appendices
%
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi
%
\bibliographystyle{plain}
\bibliography{reference}
\end{document}

>>>>

\documentclass[journal]{IEEEtran}
\usepackage{blindtext}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{pinyin}
\usepackage{natbib}
\usepackage{notoccite}
\usepackage{amssymb}
\usepackage{textcomp}
\usepackage{floatrow}
%
\begin{document}
%
% paper title
\title{Real-Time Application of Activity Classification from Body-Worn Accelerometers}
%
\author{Craig~Euler, C.T. Lin, Melissa Flores, et al? (Ruben, Brian, Sydney, Steve?)}
%
% make the title area
\maketitle
%
\begin{abstract}
Body worn sensors have been used in the market for many years now.
Many of which are used for activity monitoring or fitness.
In this paper, we make use of a matched-filtering algorithm for real-time activity classification from data obtained with three-axis accelerometers worn by the user.
We achieve invariance to sensor orientation with dimensionality reduction.
We make use of an instance-based learning algorithm to train the device to learn the individual's motion patterns and store that information as activity filters for our matched filter.
We improve computational efficiency with dimensionality reduction, temporal decomposition, and pre-processing of our filters for later use.
An assessment of our computational costs are provided as processing flop counts for the real-time specific parts of the algorithms.
\end{abstract}
%
\section{Introduction}
Various methods for activity classification exist.
Custom dicision tree, automatically generated decision tree, and artificial neural networks \cite{parkka_ermes_korpipaa_mantyjarvi_peltola_korhonen_2006} have been explored as well as other frequency based classification methods \cite{sharma_purwar_lee_lee_chung_2008}.
In this paper, we employ the use of a matched-filter method to identify an individual’s activity in real-time.
The matched-filtering method with use of thresholding and comparing the correlations of various filters is a simple method to use and have been studied in the past \cite{giannakis_tsatsanis_1990}.
Computational resources are limited on micro-processors so we've designed our algorithms to be as efficient as possible.
Having an a priori knowledge of the activity the individual is performing is useful for many current fitness algorithms.
These devices (Fit-Bit, Apple watch, Moov, Jawbone, etc.) usually have the individual input the activity they are performing.
%
\section{Activity Classification via Matched Filtering}
The matched-filter method is commonly used for extracting a signal out of noisy data.
We employ this method for identifying a user’s activity from data generated by body-worn accelerometers.
Because the matched-filter method requires use of an existing reference template to identify the desired signal, we must develop our template from a training set.
We will discuss how we develop our set of templates for use in our matched-filter.
%
\section{Dimensionality Reduction}
We manage to reduce our computational resources and remove our dependence on sensor orientation at the same time through a process of dimensionality reduction.
We assume that over a short period of time a user’s limb movement is constrained mostly to a two-dimensional plane.
By identifying the plane of motion, we then project the acceleration data onto this plane and re-define our $x$ and $y$ axis accordingly.
By this assumption, activities that are mostly confined to two dimensions will improve our signal identification.

@@

The training module, like the pre-processing module, does not need to be in real time.
This algorithm is the most demanding on computational resources and is rarely ran so this can be performed on a smart phone or tablet.
The purpose of the training module is to derive the reference signal to be used on the micro-controller for real-time processing.
The training process uses the same algorithm kernels as the activity classification.
%
\section{Results}
%
ROC curves, timing results?, flop counts.
\begin{figure}[!tbp]
\RawFloats
  \centering
     \begin{minipage}[b]{0.45\textwidth}
        \includegraphics[angle=0,width=\textwidth]{ROC_walk_to_run.png}
        \caption{\label{ROC_walk_to_run_fig}ROC curve with a downsample factor of 8}
     \end{minipage}
     \hfill
     \begin{minipage}[b]{0.45\textwidth}
     \includegraphics[angle=0,width=\textwidth]{ROC_walk_to_bike.png}
        \caption{\label{ROC_walk_to_bike_fig}ROC curve with a downsample factor of 8}
     \end{minipage}
\end{figure}
%\FloatBarrier
reference to figure \ref{ROC_walk_to_run_fig}
%
\section{Conclusion}
conclusion text.
%
\appendices
%
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi
%
\bibliographystyle{plain}
\bibliography{reference}
\end{document}
